\chapter{Introduction}

Mathematical optimization is concerned with problems of the form
\[
  \min_{x\in \mathcal{X}} \, f(x),
\]
where $f\colon\R^n\to\R\cup\{\infty\}$ is a function and
$\mathcal{X}\subset\dom(f)$ is a set.
Typically, the set $\mathcal{X}$ is called the \textit{feasible set},
elements $x\in\mathcal{X}$ are called \textit{feasible points}, and $f$ is called the \textit{objective function},
or simply \textit{objective}.
Oftentimes, one refers to $x\in\R^n$ as a \textit{decision variable}.
If $\mathcal{X}$ is nonempty, then the problem is called \textit{feasible}.
In that case,
if there exists $x^\star\in\mathcal{X}$ such that $f(x^\star) \leq f(x)$ for all $x\in\mathcal{X}$,
then $x^\star$ is called \textit{solution}. The value $f(x^\star)$ is called
\textit{optimal value} or \textit{minimal value}.
In practice, the feasible set $\mathcal{X}$ is often defined implicitly through the use of auxilliary functions,
which leads to the formulation
\begin{equation}
    \label{eq:general-constrained-problem}
    \begin{aligned}
    &\min_{x \in \mathbb{R}^n} \, f(x) \\
    &\textup{subject to (s.\,t.)} \enspace &g(x) \leq 0 \notag \\
    &\phantom{\textup{s.\,t.}} \enspace &h(x) = 0 \notag
    \end{aligned}
\end{equation}
for functions $g\colon\R^n\to\R^m$, $h\colon\R^n\to\R^\ell$. The feasible
set is then given by $\mathcal{X} = \{\, x\in\R^n \mid g(x) \leq 0 \text{ and } h(x) = 0 \,\}$.
An important special case of optimization problems are
\textit{convex optimization problems}, where the objective $f$ and
the map $g$ are convex functions, and $h$ is affine.
Convex problems have (among other things) the highly desirable property that
every local minimum is a global minimum, which makes
gradient-based optimization algorithms
work effectively \cite{boyd2004convex}.

In this work, we will consider
\textit{stochastic} convex optimization problems, which incorporate uncertainty
into the problem formulation through the use of random variables
\cite{robbins1951stochastic,birge2011introduction}.
Such problems appear in many areas of
applied mathematics, including machine learning \cite{bottou2018optimization},
supply chain management \cite{santoso2005stochastic}, and
portfolio optimization \cite{mulvey1995robust}.
Specifically, the stochastic optimization problems that we will consider
take the form
\begin{equation}
    \label{eq:main-problem}
    \tag{P}
    \begin{aligned}
    &\min_{x \in \mathbb{R}^n} \, \{\, f(x) := \E(F_\xi(x)) + r(x) \,\} \\
    &\textup{subject to} \quad g(x,\xi) \leq 0 \quad \textnormal{almost surely},
    \end{aligned}
\end{equation}
where $\xi$ is a random variable that captures uncertainty in the objective and
the constraints.
The objective $f$ is composed of two functions: the expectation
functional $\E(F_\xi)\colon\R^n\to\R$, which we assume to
be smooth, and the potentially nonsmooth function $r\colon\R^n\to\R\cup\{\infty\}$,
which is sometimes referred to as the \textit{regularizer}.
Notable instances of problems that can be formulated in this way
include primal support vector machines \cite{fercoq2019almost,hastie2009elements}, and
robust optimization by sampling \cite{ben2009robust,calafiore2005uncertain,xu2020primal}.
We are particularly interested in the case where the number of constraints
induced by the distribution of $\xi$ is large, possibly even infinite.
These \textit{large-scale} constraints prevent the use of classical methods
in deterministic optimization, which need to process all constraints at once.
This motivates the development of a method that only needs to process
a small amount of constraints at a time.

% Closely related to problems like \eqref{eq:main-problem-copy} are \textit{chance constrained}
% optimization problems, which take the form
% \begin{equation*}
%     \begin{aligned}
%     &\min_{x \in \mathbb{R}^n} \, \{\, f(x) := \E(F_\xi(x)) + r(x) \,\} \\
%     &\textup{subject to} \quad \Prob(A(\xi) x - b(\xi) \leq 0) \geq 1 - \alpha
%     \end{aligned}
% \end{equation*}
% for some $\alpha \in [0, 1)$. Chance constrained problems are
% difficult for several reasons, one being that
% the constraints are nonconvex for most probability measures $\Prob$.

A popular approach to solve \eqref{eq:main-problem} is an iterative method called \textit{stochastic gradient
descent} (SGD), introduced by Robbins and Monro \cite{robbins1951stochastic}.
SGD is very similar to the classic \textit{gradient descent} algorithm \cite{boyd2004convex,nocedal2006numerical}.
The difference is that the gradients are replaced
by \textit{stochastic gradients}: A stochastic gradient of $f$ at $x\in\R^n$ is a
random variable $g$ such that $\E(g) = \nabla f(x)$. For example, if $f(x) = \E(F_\xi(x))$ and $F_\xi$
satisfies certain regularity conditions, then
$\nabla f(x) = \E(\nabla F_\xi(x))$, which implies that $\nabla F_\xi(x)$ is a stochastic gradient of $f$
at $x$ for every $x\in\R^n$.
The advantage is that the computation of the gradient $\nabla f(x)$ is replaced with the (potentially much) cheaper
computation of $\nabla F_\xi(x)$.

Unfortunately, classic SGD is not generally applicable to problems such as \eqref{eq:main-problem},
because in order to deal with the constraints, the algorithm
needs to compute the \textit{projection operator}, $x\mapsto \argmin_{y\in \mathcal{X}} \norm{x - y}$
with $\norm{\cdot}$ denoting the euclidian norm on $\R^n$,
in every iteration.
This computation is well known to
be tractable only for certain (simple) kinds of constraints \cite{duchi2018introductory}.
In contrast, problem \eqref{eq:main-problem} may have a highly complex set of constraints, possibly even infinitely many.

A standard method for dealing with
constraints in
% deterministicinequality-constrained
deterministic constrained optimization problems like
\eqref{eq:general-constrained-problem}
% with $h\equiv 0$,
involves the use of \textit{penalty functions}
$\pi\colon\R^n\to[0,\infty)$ such that $\pi(x) = 0$ for all $x\in\mathcal{X}$ and $\pi(x) > 0$
otherwise.
Instead of the constrained problem \eqref{eq:general-constrained-problem}, one considers
the unconstrained problem
\[
    \min_{x\in\R^n} \, f(x) + \gamma \pi(x),
\]
where $\gamma\in(0,\infty)$ is a parameter that controls the degree to which infeasible points get penalized.
In the case where the problem is purely inequality constrained, that is $h\equiv 0$,
popular examples of penalty functions are
\begin{itemize}
    \item[-] the nonsmooth penalty $\pi_{\ell_1}(x) := \norm{(g(x))_+}_1$ with $\norm{\cdot}_1$ the $\ell_1$-norm on $\R^m$, and
    \item[-] the smooth penalty $\pi_{\ell_2}(x) := \norm{(g(x))_+}_2^2$ with $\norm{\cdot}_2$ the $\ell_2$-norm on $\R^m$,
\end{itemize}
where we denote by $(y)_+$ the elementwise application of $x\mapsto \max(0, x)$ to the vector $y\in\R^m$.
Unfortunately, these penalty functions suffer from well-known problems: The smooth penalty $\pi_{\ell_2}$ will, in general, not
recover a feasible point of \eqref{eq:general-constrained-problem} --- no matter how large $\gamma$
is chosen. To get reasonably close to a feasible point, one is often forced to
choose a very large $\gamma$, which introduces issues related to poor conditioning. On the other hand, the
nonsmooth penalty $\pi_{\ell_1}$ is \textit{exact}, which means that there does
exist a penalty parameter $\gamma^\star\in(0,\infty)$ such that the minimizer
of $f + \gamma^\star \pi_{\ell_1}$ is the same as that of the original constrained problem.
However, it is not clear in general how to choose $\gamma^\star$, and
the non-differentiability introduces problems for gradient-based methods \cite{nocedal2006numerical,bertsekas1997nonlinear}.
Lastly, as these penalties need to evaluate all $m$ constraints at once, they are not applicable
to the large-scale problems that we care about.
% Penalty methods are typically implemented with a two-loop iterative algorithm, where, in the inner loop,
% one applies some optimization method to \dots for some number of iterations.
% Then, in the outer loop, $\gamma$ is increased, and one again enters the inner loop, this time starting
% from the iterate found through the previous inner loop run. This process is repeated for some fixed number of iterations,
% or until some stopping criterion is met.

An alternative to the aforementioned penalty functions, which avoids many of their drawbacks,
was introduced by Nedić et al. \cite{nedich2023huber}. In their work,
the authors analyzed a special case of our problem \eqref{eq:main-problem}, namely
\begin{equation}
    \label{eq:nedic-problem}
    \begin{aligned}
    &\min_{x \in \R^n} \, f(x) \\
    &\textup{s.\,t.} \quad a_i^\top x - b_i \leq 0 \quad \textnormal{for all $i\in\{\, 1,\dots,m \,\}$}.
    \end{aligned}
\end{equation}
Their method to solve \eqref{eq:nedic-problem} is a penalty method. However,
instead of a fixed penalty function $\pi$, the authors
introduced a sequence of smooth penalty functions $(\pi^{\textnormal{hub}}_k)_{k\in\N}$,
defined as follows: Let $(\delta_k)_{k\in\N}$ be a sequence of positive real numbers. For $k\in\N$, we define
\[
    \pi_k^\textnormal{hub}(x) := \frac{1}{m} \sum_{i=1}^{m} h_k^{\text{hub}}(x; a_i, b_i),
\]
where
\[
    h_k^{\text{hub}}(x; a, b)
    :=
    \begin{cases}
        \frac{\langle a, x \rangle - b}{|\!|a|\!|} \quad&\textnormal{if $\langle a, x \rangle - b > \delta_k$}, \\
        \frac{(\langle a, x \rangle - b + \delta_k)^2}{4\delta_k |\!|a|\!|} \quad&\textnormal{if $-\delta_k \leq \langle a, x \rangle - b \leq \delta_k$}, \\
        0, \quad&\textnormal{if $\langle a, x \rangle - b < -\delta_k$},
    \end{cases}
\]
for $x,a \in\R^n$, and $b\in\R$. The crucial properties of the sequence $(\pi_k^\text{hub})_{k\in\N_0}$
are that (i) it is a sequence of smooth functions, (ii) the sequence converges to the exact penalty function $x\mapsto\norm{\max(0, x)}_1$
as $k\to\infty$, and (iii) the gradients of $(\pi_k^\text{hub})_{k\in\N_0}$ are uniformly bounded, in the sense that
there exists a constant $G\in(0,\infty)$ such that
$\nabla \pi_k^\text{hub}(x) \leq G$ for all $x\in\R^n$ and $k\in\N_0$.
These properties are very desirable for mathematical analysis.
Indeed,
the authors designed a single-loop iterative method around the penalty sequence $(\pi_k^\text{hub})$,
which only requires the evaluation of one constraint per iteration. They proved that
--- under reasonable assumptions --- their method converges to the solution of problem \eqref{eq:nedic-problem}.
More specifically, for any $\epsilon\in(0,\infty)$,
one can achieve convergence rates of $\mathcal{O}(\log^{(1+3\epsilon)/2} / \sqrt{K})$ in the convex case, and
$\mathcal{O}(\log^{\epsilon} / K)$ in the case where the objective is \textit{strongly convex},
as the number of iterations, $K\in\N$, goes to infinity.
This almost matches the rates of $\mathcal{O}(1/\sqrt{K})$ and $\mathcal{O}(1/K)$
achievable by SGD in the convex and strongly convex settings, respectively, while being projection-free \cite{robbins1971convergence,nemirovski2009robust}.
It should also be highlighted that the method is single-loop, as opposed to the typical
two-loop design of the previously mentioned penalty methods,
which makes for a particularly simple implementation.

Apart from \cite{nedich2023huber},
among existing work on constrained stochastic optimization,
the settings considered by Fercoq et al. \cite{fercoq2019almost},
Li et al. \cite{li2023softplus}, and Xu et al. \cite{xu2020primal}
are the most closely related to ours.
Of those, \cite{fercoq2019almost} and \cite{nedich2023huber} consider
the special case of affine constraints, and in \cite{xu2020primal},
\cite{nedich2023huber}, and \cite{li2023softplus}, the authors
assume a finite constraint set. The setting we consider therefore generalizes those of prior works.
Further, \cite{li2023softplus} focuses on strongly convex finite-sum objectives, in order
to leverage accelerated stochastic gradient methods such as SVRG \cite{johnson2013svrg}.
In contrast, we will consider objectives that contain general expected value functions,
and provide analysis for both strongly convex, as well as general convex objectives.
Finally, both \cite{fercoq2019almost} and \cite{xu2020primal}
make assumptions to ensure that the stochastic
gradients have uniformly bounded variance. However, this assumption does not apply to many
realistic settings \cite{nguyen2018sgd} and we will not rely on it for the
analysis of our method.

Our contributions in this work are as follows:
We build on the ideas of Nedić et al. \cite{nedich2023huber} with a new
algorithm that exploits the compositional structure in the objective and
applies to a wider range of constraints.
Specifically, our method and the analysis thereof apply to the
general setting of problem \eqref{eq:main-problem},
which allows for infinitely many nonlinear constraints and
makes no finite-sum assumption on the objective. This clearly differentiates
it from the settings considered in prior works.
A notable feature of our proofs is that we do not make any global boundedness assumptions on the norm
of the stochastic gradients, which stands in contrast to many related works in the stochastic optimization
literature.

In the following, we will first begin with a review of the relevant background material in the
fields of probability and convex optimization in \cref{chap:background}.
The main contributions of this work
are then presented in \cref{chap:sipm}, where we formulate our algorithm
and prove various convergence results under relatively mild assumptions on the problem setting beyond convexity.
We support our theoretical analysis with numerical experiments in \cref{chap:numerical-experiments}.
Finally, we summarize our findings and point to future research directions in \cref{chap:outlook}.